---
title: 'Problem Set #4'
author: "Seeam S. Noor"
date: "Due February 27, 2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# Loading all packages necessary to complete this assignment
library(gt)
library(tidyverse)
library(lubridate)
library(forcats)
library(readr)

# Creating the data frame from the csv file in project directory
orig <- read_csv(
  file = "ps_4_elections-poll-nc09-3.csv",
  # Making sure each column is read as the right data type
  col_types = cols(
    .default = col_character(),
    turnout_scale = col_double(),
    turnout_score = col_double(),
    w_LV = col_double(),
    w_RV = col_double(),
    final_weight = col_double(),
    timestamp = col_datetime(format = "")))
```


```{r question1, echo=FALSE, message=FALSE, warning=FALSE}
  # Filtering tibble for only Democrat voters and counting them
  Q1a <- orig %>% filter(response == 'Dem') %>% nrow()
  
  # Creating a new dataframe that will help answer the 2nd question
  Q1b_table <- 
    orig %>% 
    # Filtering Undecided and Republicans and then grouping by that and counting them
    filter(response %in% c('Rep','Und')) %>% 
    group_by(response) %>% 
    count()
  
  # Extracting the number in each category and calculating the difference
  Q1b <- Q1b_table$n[1] - Q1b_table$n[2]
  
  Q1c <- 
    orig %>% 
    # Choosing the columns we need
    select(gender, gender_combined) %>%
    # Creating a new column that gives 1 if our selected columns, 0 otherwise
    mutate(compare = if_else(gender == gender_combined, 1, 0)) %>%
    # Filtering the number of 0s and counting the number of rows we get
    filter(compare == 0) %>%
    nrow()
  
  Q1d <- 
    orig %>% 
    # Choosing the columns we need
    select(race_eth, file_race_black) %>%
    # Filtering for only the rows we need
    filter(race_eth == 'White') %>%
    # Creating a new column that gives 1 if our selected columns, 0 otherwise
    mutate(compare = if_else(race_eth == file_race_black, 1, 0)) %>%
    # Filtering the number of 0s and counting the number of rows we get
    filter(compare == 0) %>%
    nrow()
  
  Q1e_data <-
    orig %>%
    # Filtering for only Democrats and Republicans
    filter(response %in% c('Dem','Rep')) %>%
    # Choosing the columns we need
    select(response, timestamp) %>%
    # Grouping by Democrats and REpublicans
    group_by(response) %>%
    # Taking the timestamp which happened earliest in both groups
    summarize(fastest = min(timestamp))
  
  # Taking the difference in minutes between the two stamps
  Q1e <- minute(Q1e_data$fastest[2]) - minute(Q1e_data$fastest[1])
  

  # Answers to all questions have been saved as separate variables so they can be easily included in the text below
```

## Question 1
  
There were `r Q1a` respondents who supported the Democratic candidate.  
There were `r Q1b` more respondents who favored the Republican candidate than who were Undecided.  
There are two gender variables (gender and gender_combined). There are `r Q1c` individuals for whom these variables have different values.  
There are `r Q1d` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.  
The first response of Dem came `r Q1e` minutes (rounded to the nearest minute) before the first response of Rep.

##Question 2  
  

```{r question2, echo=FALSE, message=FALSE, warning=FALSE}
  # Creating a new dataframe for our plot
  Q2_data <- 
     orig %>% 
      # Choosing the columns we need
      select(response, race_eth, final_weight) %>%
      # Removing people who didn't respond
      filter(race_eth != "[DO NOT READ] Don't know/Refused") %>%
      # Making sure our categories are in right order
      mutate(race_eth = fct_relevel(race_eth, c("White", "Black", "Hispanic", "Asian","Other"))) %>%
      # Grouping by response and race
      group_by(response, race_eth) %>%
      # Considering the fact that some respondents are weighted more since they respond less
      summarize(total = sum(final_weight)) %>%
      # splitting the data on response as key 
      spread(key =  response, value = total)
     
  # Turning all NAs into 0
  Q2_data[is.na(Q2_data)] <- 0   
     
  Q2_data %>%
    # Creating a column that is sum of Rep, Dem and Und
    mutate(all = Dem + Rep + Und + 3) %>%
    # Making sure the columns are proportions of total
    mutate(Dem = Dem / all) %>%
    mutate(Rep = Rep / all) %>%
    mutate(Und = Und / all) %>%
    # Choosing the columns we need
    select('race_eth','Dem','Rep','Und') %>%
    # Renaming the race_eth column to something easier to understand
    rename(Race = race_eth) %>%
    # Using gt to create a better looking table
    gt() %>%
    # Adding a header to the table
    tab_header(
      title = "Polling Results in North Carolina 9th Congressional District by Race") %>%
    # Fixing the labels for the table
    cols_label(
      Dem = "DEM.",
      Rep = "REP.",
      Und = "UND."
    ) %>%
    # Turning the numbers into rounded percentages
    fmt_percent(columns = vars(Dem, Rep, Und),
                 decimals = 0) %>%
    # Turning the 0s into NAs
    na_if(0) %>%
    # Reformatting the NAs into dashes
    fmt_missing(columns = vars(Und), rows = 4)
```

##Question 3  


```{r question3, echo=FALSE, message=FALSE, warning=FALSE}
  orig %>%
    # Filtering people who didn't want to respond
    filter(educ != '[DO NOT READ] Refused') %>%
    # Choosing the columns we need
    select(educ, final_weight) %>%
    group_by(educ) %>%
    # Plotting our data and arranging the educational levels by weight given
    ggplot(aes(x = fct_reorder(educ,desc(final_weight)), y = final_weight)) +
    # Creating a violin plot
    geom_violin() +
    # Flipping the plot to make it a horizontal violin plot
    coord_flip() +
    # Plotting the individual points in tnside each violin
    geom_jitter(height = 0, width = 0.1) + 
    # Removing the x-axis label(y-axis practically since we flipped the plot)
    xlab(NULL) + 
    # Providing appropriate y-axis title
    ylab("Weight Given to Respondents in Calculating Poll Results") +
    # Adding title, subtitle and caption to the plot
    labs(title = "More Educated Matter Less in North Carolina 9th", subtitle = "Poll gives more weight to people who are less likely to participate in polls", caption = "New York Times Upshot/Sienna College 2018 live polls")
```

##Question 4  



```{r question4, echo=FALSE, message=FALSE, warning=FALSE}
  orig %>% 
    # Choosing the columns we need
    select(ager, gender, response) %>%
    # Filtering for responds we don't need
    filter(ager != "[DO NOT READ] Refused" & response != '3' & response != 'Und') %>%
    # Grouping data by response, age and gender
    group_by(response, ager, gender) %>% 
    # Choosing aesthetics where gender as fill
    ggplot(aes(x = ager, fill = gender)) +
    # Creating a bar graph
    geom_bar() +
    # Creating two seperate graphs, Rep and Dem, to see genderwise response in each age category
    facet_wrap(~response) + 
    # Flipping the axes to get horizontal bars
    coord_flip() + 
    # Removing the x-axis title(y-axis practically since we flipped the plot)
    xlab(NULL) + 
    # Providing appropriate y-axis title
    ylab("Number of respondents in each category") +
    labs(title = "Party preferences by gender and age", subtitle = "Females have more Democrats in most age categories \nMales have more Republicans in each age category", caption = "New York Times Upshot/Sienna College 2018 live polls")
```



##Colleagues
  
Christopher Onesti